### 2025-12-14 extra/paper-quollr-code.Rmd
Rendered new paper quollr version

---
title: "paper-quollr-code"
format: pdf
---

```{r, fig.cap="Line plot."}
#| label: fig-unemployment

library(autoalt)
library(ggplot2)
library(dplyr)

# Built-in datasets from ggplot2 and base R
data("economics", package = "ggplot2")
data("airquality")

economics |>
  ggplot(aes(x = date, y = unemploy)) +
  geom_line(linewidth = 0.6) +
  labs(
    title = "Total Number of Unemployed Persons in the US Over Time",
    x = "Year",
    y = "Number of Unemployed (thousands)"
  ) +
  theme_minimal()

```