### 2025-12-14 extra/paper-quollr-code.Rmd
Fixed the issues

---
title: "Sample quollr"
author: "Chat-GPT"
date: "2025-09-10"
format: pdf
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false
library(quollr)
library(tibble)
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(patchwork)
library(readr)

# Scale high-d data

# Center the data by subtracting the mean of each column
center_data <- function(data) {
  apply(data, 2, function(col) col - mean(col))
}

# Function to scale data manually
scale_data_manual <- function(data, type_col) {
  # Step 1: Center the data (mean 0)
  data_centered <- center_data(data |> select(-all_of(type_col)))

  # Step 2: Calculate the standard deviation of each dimension
  sds <- apply(data_centered, 2, sd)

  # Step 3: Scale each dimension to have the range [0, 1]
  data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))

  # Step 4: Scale dimensions according to their variation
  # The dimension with the highest standard deviation is scaled to [-1, 1]
  # Other dimensions are scaled to smaller ranges based on their standard deviations
  max_sd <- max(sds)

  # Normalize the standard deviations to get scaling factors
  scaling_factors <- sds / max_sd

  for (i in seq_along(scaling_factors)) {
    data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
  }

  # Combine the scaled data with the 'type' column and return as a tibble
  data_scaled <- as_tibble(data_scaled) %>%
    mutate(!!type_col := data[[type_col]])

  return(data_scaled)
}

# creating Standardization function
standardize = function(x){
  z <- (x - mean(x)) / sd(x)
  return( z)
}

# Solve quadratic function

quad <- function(a = 3, b = 2 * a2, c = -(a2^2 + a1^2))
{
  a <- as.complex(a)
  answer <- c((-b + sqrt(b^2 - 4 * a * c)) / (2 * a),
              (-b - sqrt(b^2 - 4 * a * c)) / (2 * a))
  if(all(Im(answer) == 0)) answer <- Re(answer)
  if(answer[1] == answer[2]) return(answer[1])
  answer[answer>0] ## only positive
}
```

```{r}
#| label: fig-overview
#| fig-cap: "Wireframe model representation of the NLDR layout, lifted and displayed in high-dimensional space. The left panel shows the NLDR layout with a triangular mesh overlay, forming the wireframe structure. This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved. Panels (a1–a4) display different 2D projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible. The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions."
#| echo: false
# Plot MSE

plot_hbe <- function(error_df) {

  ggplot(error_df,
         aes(x = a1,
             y = HBE,
             colour = method)) +
    geom_point(size = 0.8) +
    geom_line(linewidth = 0.3) +
    ylab("HBE") +
    xlab(expression(paste("binwidth (", a[1], ")"))) +
    theme_minimal() +
    theme(panel.border = element_rect(fill = 'transparent'),
          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
          axis.ticks.x = element_line(),
          axis.ticks.y = element_line(),
          legend.position = "none",
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.x = element_text(size = 7),
          axis.title.y = element_text(size = 7),
          plot.margin = margin(0, 0, 0, 0))

}

set.seed(20240110)

clr_choice <- "#66B2CC"

scurve_model_obj <- fit_highd_model(
  highd_data = scurve, 
  nldr_data = scurve_umap, 
  b1 = 21, 
  q = 0.1, 
  hd_thresh = 0)

scurve_umap_scaled <- scurve_model_obj$nldr_scaled_obj$scaled_nldr
tr_from_to_df_scurve <- scurve_model_obj$trimesh_data
df_bin_centroids_scurve <- scurve_model_obj$model_2d
df_bin_scurve <- scurve_model_obj$model_highd
hex_grid_scurve <- scurve_model_obj$hb_obj$hex_poly
counts_df_scurve <- scurve_model_obj$hb_obj$std_cts

hex_grid_with_counts_scurve <- left_join(
  hex_grid_scurve, counts_df_scurve, 
  by = c("h" = "h"))

hex_grid_nonempty_scurve <- hex_grid_scurve |>
  filter(h %in% df_bin_centroids_scurve$h)

sc_xlims <- c(-0.25, 1.35)
sc_ylims <- c(-0.25, 1.35)

scurve_umap_plt <- ggplot(
  scurve_umap_scaled, 
  aes(x = emb1, y = emb2)) +
  geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
  xlim(sc_xlims) +
  ylim(sc_ylims)

hex_grid_poly_scurve <- ggplot(
  data = hex_grid_with_counts_scurve, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = h), 
               fill = "#ffffff") +
  geom_point(data = scurve_umap_scaled, 
             aes(x = emb1, y = emb2), 
             alpha = 0.8, size = 0.5, color = clr_choice) + 
  xlim(sc_xlims) +
  ylim(sc_ylims)

hex_centroids_scurve <- ggplot(
  data = hex_grid_with_counts_scurve, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = h), 
               fill = "#ffffff")  +
  geom_point(data = df_bin_centroids_scurve, 
             aes(x = c_x, y = c_y), 
             size = 1, color = "#FF7755") +
  xlim(sc_xlims) +
  ylim(sc_ylims)

wireframe_scurve <- ggplot() +
  geom_segment(data = tr_from_to_df_scurve,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = df_bin_centroids_scurve, 
             aes(x = c_x, y = c_y), 
             size = 1, color = "#FF7755") +
  xlim(sc_xlims) +
  ylim(sc_ylims)

## 2-d vis model
hex_grid_scurve <- ggplot(
  data = hex_grid_with_counts_scurve, 
  aes(x = x, y = y)) +
  geom_point(data = scurve_umap_scaled, 
             aes(x = emb1, y = emb2), 
             size = 0.5, color = clr_choice) +
  geom_segment(data = tr_from_to_df_scurve,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = df_bin_centroids_scurve, 
             aes(x = c_x, y = c_y), 
             size = 1, color = "#FF7755")

## High-d vis model
df_exe <- comb_data_model(
    highd_data = scurve, 
    model_highd = df_bin_scurve, 
    model_2d = df_bin_centroids_scurve
)

df <- df_exe |>
    dplyr::filter(type == "data") ## original dataset

df_b <- df_exe |>
  dplyr::filter(type == "model") ## High-d model

scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                         lineFrom = tr_from_to_df_scurve$from_reindexed,
                         lineTo = tr_from_to_df_scurve$to_reindexed,
                         group = df_exe$type,
                         pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
                         levelColors = c(clr_choice, "#FF7755"),
                         enableControls = FALSE,
                         width = "400px", height = "380px")

scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                         lineFrom = tr_from_to_df_scurve$from_reindexed,
                         lineTo = tr_from_to_df_scurve$to_reindexed,
                         group = df_exe$type,
                         pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
                         levelColors = c(clr_choice, "#FF7755"),
                         enableControls = FALSE,
                         width = "400px", height = "550px")

df_b_scurve <- df_bin_scurve |>
  dplyr::filter(h %in% df_bin_centroids_scurve$h) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

## Reorder the rows of df_b according to the h order in df_b_with_center_data
df_b_scurve <- df_b_scurve[match(df_bin_centroids_scurve$h, df_b_scurve$h),] |>
  dplyr::select(-h)

scurve_labeled <- scurve |>
  select(-ID) |>
  mutate(type = "data")

# Apply the scaling
df_model_data_scurve <- bind_rows(scurve_labeled, df_b_scurve)
scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
  as_tibble()

scaled_scurve_data <- scaled_scurve |>
  filter(type == "data") |>
  select(-type)

scaled_scurve_data_model <- scaled_scurve |>
  filter(type == "model") |>
  select(-type)

## First projection
projection <- cbind(
  c(0.10479,0.06673,0.19430,-0.14763,0.02861,-0.04302,0.06601),
  c(0.11421,0.18556,-0.16860,-0.05033,-0.00219,0.04159,0.04285))

proj_obj1 <- get_projection(projection = projection, 
                            proj_scale = 1.23, 
                            highd_data = scaled_scurve_data, 
                            model_highd = scaled_scurve_data_model, 
                            trimesh_data = tr_from_to_df_scurve, 
                            axis_param = list(limits = 0.5, 
                                              axis_scaled = 2,
                                              axis_pos_x = -0.25, 
                                              axis_pos_y = -0.25, 
                                              threshold = 0.0259))

scurve_proj_umap_model1 <- plot_proj(
  proj_obj = proj_obj1,
  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
  plot_limits = c(-0.35, 0.35), 
  axis_text_size = 2,
  is_category = FALSE)

scurve_proj_umap_model1_dp <- plot_proj(
  proj_obj = proj_obj1,
  point_param = c(0.2, 0, clr_choice), # size, alpha, color
  plot_limits = c(-0.35, 0.35), 
  axis_text_size = 3,
  is_category = FALSE)

## Second projection

projection <- cbind(
  c(0.01588,-0.13640,-0.17180,-0.15486,-0.09726,0.00858,0.01090),
  c(0.21443,-0.00390,0.08182,-0.11746,0.06820,-0.05458,-0.08885))

proj_obj2 <- get_projection(projection = projection, 
                            proj_scale = 1.23, 
                            highd_data = scaled_scurve_data, 
                            model_highd = scaled_scurve_data_model, 
                            trimesh_data = tr_from_to_df_scurve, 
                            axis_param = list(limits = 0.5, 
                                              axis_scaled = 2,
                                              axis_pos_x = -0.25, 
                                              axis_pos_y = -0.25, 
                                              threshold = 0.0259))

scurve_proj_umap_model2 <- plot_proj(
  proj_obj = proj_obj2,
  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
  plot_limits = c(-0.35, 0.35), 
  axis_text_size = 2,
  is_category = FALSE)

## Third projection

projection <- cbind(
  c(0.02727,0.09154,-0.16383,-0.04931,-0.17013,0.11965,-0.01884),
  c(-0.10510,-0.10971,-0.09084,0.19642,-0.07163,-0.02486,0.07965))

proj_obj3 <- get_projection(projection = projection, 
                            proj_scale = 1.23, 
                            highd_data = scaled_scurve_data, 
                            model_highd = scaled_scurve_data_model, 
                            trimesh_data = tr_from_to_df_scurve, 
                            axis_param = list(limits = 0.5, 
                                              axis_scaled = 2,
                                              axis_pos_x = -0.25, 
                                              axis_pos_y = -0.25, 
                                              threshold = 0.031))

scurve_proj_umap_model3 <- plot_proj(
  proj_obj = proj_obj3,
  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
  plot_limits = c(-0.35, 0.35), 
  axis_text_size = 2,
  is_category = FALSE)

## Fourth projection

projection <- cbind(
  c(0.17174,0.14255,0.11950,0.10344,-0.07148,-0.01656,0.04275),
  c(0.09270,0.04486,-0.14397,0.08305,0.11783,0.16187,-0.06077))

proj_obj4 <- get_projection(projection = projection, 
                            proj_scale = 1.23, 
                            highd_data = scaled_scurve_data, 
                            model_highd = scaled_scurve_data_model, 
                            trimesh_data = tr_from_to_df_scurve, 
                            axis_param = list(limits = 0.5, 
                                              axis_scaled = 2,
                                              axis_pos_x = -0.25, 
                                              axis_pos_y = -0.25, 
                                              threshold = 0.031))

scurve_proj_umap_model4 <- plot_proj(
  proj_obj = proj_obj4,
  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
  plot_limits = c(-0.35, 0.35), 
  axis_text_size = 2,
  is_category = FALSE)

hex_grid_scurve + wrap_plots(
  scurve_proj_umap_model1, scurve_proj_umap_model2,
  scurve_proj_umap_model3, scurve_proj_umap_model4, 
  ncol = 2)
```

# Introduction

This report explores two publicly available datasets: economics from the ggplot2 package and the classic airquality dataset from base R. The economics dataset provides monthly time-series data on the US economy, including variables such as unemployment and personal consumption expenditure. The airquality dataset records daily air quality measurements in New York, including ozone concentration, solar radiation, and temperature.

We use a selection of visualisations to highlight long-term economic trends and seasonal patterns in air quality.

This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data. Starting with an NLDR layout, our approach is to create a $2\text{-}D$ wireframe model representation, that can be lifted and displayed in the high-dimensional @fig-overview.